Waving and swaying

### 2014-01-27 Paper/grass.tex
before aspen

\begin{abstract}
The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. 
We find that this inflection point is located inside the grass canopy but it does not lead to Kelvin Helmholtz instability which is 
normally observed in a free shear layer. We also show that the wavelength of coherent vortices scales with the unvegetated water depth, and not with 
the shear layer thickness as predicted by a theory based on Kelvin-Helmholtz instability. Based on these results, we propose a mechanism for waving 
marine grass based on the shear instability of the flow above the canopy.

%The spontaneous waving of marine grass is though to be due to a Kelving-helmholtz instability resulting from strong shear near top of grass. On the contrary
%we find that strong velocity shear resulting from vertical discontinuity of drag due to presence of grass does not lead to Kelvin-Helmholtz instability which is 
%normally observed in free shear flows. We also show that wavelength of coherent vortices scales with the unvegetated water depth. Based on these results, 
%we propose an alternative mechanism which attributes waving of marine grass on shear instability of flow above the canopy. 
\end{abstract}
\maketitle

@@

\begin{figure}[htb!]
  \includegraphics[]{fig1}
\caption{Basic flow profile (left )and model of canopy (bottom) along with variation of shear layer (inset) }
\end{figure}
Knowing the scale of shear layer we can systematically investigate its effect on flow instability
by varying grass number density $N_g$ and pressure gradient $\frac{dP}{dx}$. Considering small perturbations $u, v, p$ about the base profile $U$ and $P$
respectively, momentum and mass balance equation expanded in first order of perturbed variable yields.
\begin{equation}
%\scriptstyle{
\begin{split}
 %u_t+Uu_x+vU_y &= \frac{1}{\rho}p_x+\frac{\mu}{\rho}\left(u_xx+u_yy \right) -2C_NdN_gUu\\
\frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}\nabla^2u-2C_{N}dN_{g}Uu\\
\frac{\del v}{\del  t}+ U\frac{\del v}{\del x} &= -\frac{1}{\rho}\frac{\del p}{\del y}+\frac{\mu}{\rho}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0\\
 %\nabla\cdot \bu &= 0
\end{split}
%}
\end{equation}
%In most of recent work on honami/monami base profile have been approximated to a plane mixing layer profile having inflection point near top of canopy, It is also argued that 
%It is instability associated with inflection of mean velocity which is responsible for coherent eddies traveling on top of canopies. In our calculation we get base velocity
%profile by hydrodynamic balance between pressure, drag and diffusive force which also results in a velocity profile with inflection point near top of canopy with additional property 
%of having been produced through equation of motion itself.\newline
%We performed stability analysis on two different kind of base flow namely (1) flow driven by constant pressure gradient between two parallel plate (2) constant pressure gradient driven 
%flow with zero shear at top. We use following scaling to non-dimensionlize equation of motion
Momentum and mass balance equation can be non-dimensionlize using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0 = \frac{dP/dxH^2}{\mu0}$.
%Using half channel height $H$ of and associated advection time H/$U_{0}$, we use following scaling to non-dimensinlize equation 
%of motion
%   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{(dP/dx)H^2}{4\mu} \]
With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance; momentum and mass balance can be combined 
into a single equation.
\begin{equation}
\scriptstyle{
Re\left[\del_t+\del_x U \right]\grad^2\psi - U_{yy}\psi_x = \grad^4\psi-\frac{\del}{\del_y}\left(\frac{2D_{drag}}{R_e}U\psi_y\right)
}
\end{equation}
where $R_{e}= \frac{\rho U_0 H}{\mu}$ is Reynolds number and $\bar{N_g} = C_N d N_g H$ is non-dimensional grass number density and  $D_{drag} = R_{e}\bar{N_{g}}$ is 
non-dimensional drag. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma}$, we get generalized orr-sommerfield equation 
which is an eigenvalue problem whose solution provides frequency and wavelength of eddies traveling over canopy.
\begin{equation}
\begin{split}
\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi &= \frac{1}{R_{e}}\left[D^2 -k^{2} \right]^2\phi +ikU_{yy}\phi \\
&-\frac{\del}{\del y}\left(\frac{2D_{drag}}{R_e}U\phi_y\right)
\end{split}
\end{equation}
Solution of generalized Orr-Somerfield equation predicts a critical Reynolds number as function of grass number density, below which there is no hydrodynamic instability (Fig(2))

% shows 
%this critical number for different grass density and grass height along with experimental threshold observed by Grizzle (1996)
\begin{figure}[htb!]
\includegraphics[scale=0.8]{Critical_Re_vs_Ng_Grizzle}
\caption{Critical Reynolds number with flow between parallel plate for different grass height as a function of grass number density}
\end{figure}
\begin{figure}[htb!]
\includegraphics[]{K_vs_shear_width}
\caption{Critical Wavenumber for different grass height as function of shear layer thickness}
\end{figure}
\begin{figure}[htb!]
\includegraphics[scale=0.34]{Observed_vs_calculated}
\caption{Comparison between experimently oberved frequency and predicted frequency}
\end{figure}

>>>>

\begin{abstract}
%The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. 
%We find that this inflection point is located inside the grass canopy but it does not lead to Kelvin Helmholtz instability which is 
%normally observed in a free shear layer. We also show that the wavelength of coherent vortices scales with the unvegetated water depth, and not with 
%the shear layer thickness as predicted by a theory based on Kelvin-Helmholtz instability. Based on these results, we propose a mechanism for waving 
%marine grass based on the shear instability of the flow above the canopy.

%The spontaneous waving of marine grass is though to be due to a Kelving-helmholtz instability resulting from strong shear near top of grass. On the contrary
%we find that strong velocity shear resulting from vertical discontinuity of drag due to presence of grass does not lead to Kelvin-Helmholtz instability which is 
%normally observed in free shear flows. We also show that wavelength of coherent vortices scales with the unvegetated water depth. Based on these results, 
%we propose an alternative mechanism which attributes waving of marine grass on shear instability of flow above the canopy.

The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. We find that this inflection point
is located near the tip of grass canopy. Our analysis shows that flow in presence of grass become unstable either through a mechanism of Kelvin-Helmholtz instability associated with 
inflection point or via shear instability of flow above the grass.

\end{abstract}
\maketitle

@@

\begin{figure}[htb!]
  \includegraphics[scale=0.8]{fig1}
\caption{Basic flow profile (left )and model of canopy (bottom) along with variation of shear layer (inset) }
\end{figure}
Knowing the scale of shear layer we can systematically investigate its effect on flow instability
by varying grass number density $N_g$ and pressure gradient $\frac{dP}{dx}$. Considering small perturbations $u, v, p$ about the base profile $U$ and $P$
respectively, momentum and mass balance equation expanded in first order of perturbed variable yields.
%\scriptstyle{
\begin{split}
 %u_t+Uu_x+vU_y &= \frac{1}{\rho}p_x+\frac{\mu}{\rho}\left(u_xx+u_yy \right) -2C_NdN_gUu\\
\frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}\nabla^2u-2C_{N}dN_{g}Uu\\
\frac{\del v}{\del  t}+ U\frac{\del v}{\del x} &= -\frac{1}{\rho}\frac{\del p}{\del y}+\frac{\mu}{\rho}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0\\
 %\nabla\cdot \bu &= 0
\end{split}
%}
%In most of recent work on honami/monami base profile have been approximated to a plane mixing layer profile having inflection point near top of canopy, It is also argued that 
%It is instability associated with inflection of mean velocity which is responsible for coherent eddies traveling on top of canopies. In our calculation we get base velocity
%profile by hydrodynamic balance between pressure, drag and diffusive force which also results in a velocity profile with inflection point near top of canopy with additional property 
%of having been produced through equation of motion itself.\newline
%We performed stability analysis on two different kind of base flow namely (1) flow driven by constant pressure gradient between two parallel plate (2) constant pressure gradient driven 
%flow with zero shear at top. We use following scaling to non-dimensionlize equation of motion
Momentum and mass balance equation can be non-dimensionlize using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0 = \frac{dP/dxH^2}{\mu0}$.
%Using half channel height $H$ of and associated advection time H/$U_{0}$, we use following scaling to non-dimensinlize equation 
%of motion
%   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{(dP/dx)H^2}{4\mu} \]
With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance; momentum and mass balance can be combined 
into a single equation.
\scriptstyle{
Re\left[\del_t+\del_x U \right]\grad^2\psi - U_{yy}\psi_x = \grad^4\psi-\frac{\del}{\del_y}\left(\frac{2D_{drag}}{R_e}U\psi_y\right)
}
where $R_{e}= \frac{\rho U_0 H}{\mu}$ is Reynolds number and $\bar{N_g} = C_N d N_g H$ is non-dimensional grass number density and  $D_{drag} = R_{e}\bar{N_{g}}$ is 
non-dimensional drag. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma}$, we get generalized orr-sommerfield equation 
which is an eigenvalue problem whose solution provides frequency and wavelength of eddies traveling over canopy.
\begin{split}
\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi &= \frac{1}{R_{e}}\left[D^2 -k^{2} \right]^2\phi +ikU_{yy}\phi \\
&-\frac{\del}{\del y}\left(\frac{2D_{drag}}{R_e}U\phi_y\right)
\end{split}
Solution of generalized Orr-Somerfield equation predicts a critical Reynolds number as function of grass number density, below which there is no hydrodynamic instability (Fig(2)).
Variation of critical wave number with shear layer thickness for different grass height is shown in Fig-3.
% shows 
%this critical number for different grass density and grass height along with experimental threshold observed by Grizzle (1996)
\begin{figure}[htb!]
\includegraphics[]{Critical_Re_vs_delta}
\caption{Critical Reynolds number with flow for different grass height as a function of inverse shear thickness}
\end{figure}
The following features can be noted from Fig(3). For the certain grass height region of linear variation of wavenumber with inverse shear layer thickness suggest that the flow is subject 
to Kelvin-Helmholtz instability above a threshold shear thickness. Region of constant wave number suggest that length scale of the instability is no longer governed by shear layer
thickness in this this regime, rather it is governed by flow above the grass, which can be understood by asymptotic analysis in regime of small shear layer thickness and large Reynolds 
number
\begin{figure}[htb!]
\includegraphics[]{K_vs_shear_width}
\caption{Critical Wavenumber for different grass height as function of shear layer thickness}
\end{figure}
%\newline
Collapse of $\phi$ for different shear thickness in the regime of constant critical wavenumber $k$ suggest that we have obtained the asymptotic solution in limit of small shear layer 
thickness. Noting that non-dimensional flow velocity $U \sim \delta^{3/2}$ inside the grass and $R_{e} \gg 1$, we can simplify the equation-(4) considerably both with in the grass and 
above the grass canopy. Inside the grass equation simplifies into $\sigma\left(D^2-k^2\right)\phi = -\frac{\del}{\del y}\left(\frac{2\phi_y}{\delta^{3/2}R_e}\right)$. Together with simplified
force balance inside the grass canopy and constant value of $k\approx O(1)$, we can estimate that critical Reynolds number $R_e \sim \delta^{-3/2}$ for small shear layer thickness Fig(2) 
\begin{figure}[htb!]
 \includegraphics[]{Asymptotic_eig_set3}
\caption{plot of $|\phi|$ for different shear layer thickness in limit of small shear layer thickness}
\end{figure}

\begin{figure}[htb!]
\includegraphics[scale=0.34]{Observed_vs_calculated}
\caption{Comparison between experimently oberved frequency and predicted frequency}
\end{figure}
